Attribute fusion in a latent process model for time series of graphs

### 2011-05-18 ssp2011.tex
Revise introduction

\documentclass[final]{IEEEtran}
\usepackage[utopia]{mathdesign}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{mathrsfs}
\usepackage{bm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[noadjust]{cite}
%\usepackage[colon,sort&compress]{natbib}
%\numberwithin{equation}{section}
\renewcommand\arraystretch{1.2}
\let\underbrace\LaTeXunderbrace
\let\overbrace\LaTeXoverbrace
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathllapinternal#1#2{%
\llap{$\mathsurround=0pt#1{#2}$}% $
}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathclapinternal#1#2{%
\clap{$\mathsurround=0pt#1{#2}$}%
}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathrlapinternal#1#2{%
\rlap{$\mathsurround=0pt#1{#2}$}% $
}
\bibliographystyle{IEEEtran}
\begin{document}
\title{Attribute fusion in a latent process model for time series of
  graphs}
\author{Carey~E.~Priebe, Nam~H.~Lee, Youngser~Park, and Minh~Tang}%
%\thanks{Johns Hopkins University \\ Department of Applied Mathematics
%  and Statistics \\ Baltimore, Maryland 21218-2682 USA}%
\maketitle
\begin{abstract}
 We consider the problem of anomaly/change point detection given a
 time series of graphs with categorical attributes on the
 edges. Various attributed graph invariants are considered, and their
 power for detection as a function of a linear fusion parameter is
 presented.  
\end{abstract}
\begin{IEEEkeywords}
  Anomaly detection, Attributed Random Graphs, Fusion, Random Dot
  Product Graphs
\end{IEEEkeywords}
\section{Introduction}

@@

originally presented in

@@

      Q^{(i)}(t) \equiv \mathbf{Q}_0

@@

     \bar{X}_{i}(t) \equiv \bar{\pi}_0.

@@

     \bar{X}_{i}(t) \equiv \bar{\pi}_1 \quad \text{for $i \leq m$} \\
     \bar{X}_{i}(t) \equiv \bar{\pi}_0 \quad \text{for $i > m$} 

@@

 where $\bar{\pi}_0$ and $\bar{\pi}_1$ are sub-probability vectors

@@

  \pi_0, \pi_1, Q_0, Q_1)$. Define $\mathbf{Z}_0$ and $\mathbf{Z}_1$

@@

  for the continuous-time Markov chain on $\mathscr{K}$ with intensity

@@

    \Sigma_0 = \mathrm{diag}(\pi_0) \mathbf{Z}_0 + \mathbf{Z}_0^{T}

@@

    \Sigma_1 = \mathrm{diag}(\pi_1) \mathbf{Z}_1 + \mathbf{Z}_1^{T}

@@

  A second order approximation $\hat{\mathscr{V}}$ of $\mathscr{V}$
  is the sequence $\{\hat{X}(t), \hat{G}(t)\}_{t=1}^{\infty}$ where

@@

  \item For each $t$ and each $i \in [n]$, $\hat{X}^{i}(t)$ is a

@@

    vector $Z_{i}(t)$ to $\mathscr{S}$.

@@

      \mathbb{E}[Z_i(t)] \equiv \bar{\pi}_0 \\
      \mathrm{Var}[Z_i(t)] \equiv \hat{\Sigma^{0}}

@@

      \mathbb{E}[Z_i(t)] \equiv \bar{\pi}_0, \quad \text{for $i \leq

@@

      \mathbb{E}[Z_i(t)] \equiv \bar{\pi}_1, \quad \text{for $i > 

@@

      \mathrm{Var}[Z_i(t)] \equiv \hat{\Sigma}_{0} \quad \text{for $i

@@

      \mathrm{Var}[Z_i(t)] \equiv \hat{\Sigma}_{1} \quad \text{for $i

@@

    where $\hat{\Sigma}_0$ and $\hat{\Sigma}_1$ are the matrices
    obtained by removing last row and column of $\Sigma_0$ and
    $\Sigma_1$. 

@@

@@

for $l \geq 2$. 

@@

  \label{lem:1}

@@

$\mu_{\lambda}$. Theorem \ref{thm:1} extends a result in

@@

  \label{thm:1}
  Let $\zeta$ and $\xi$ be as defined in Lemma \ref{lem:1}.  Suppose

@@

@@

$T_{\lambda}^{l}(t)$ is, once again, a Student t-distribution with

@@

  \label{lem:2}

@@

likely to be $\tfrac{\rho_\lambda}{\varsigma_\lambda}$. For sufficiently large $n$
and $m$, we have

@@

    \sigma_A} \\ &= \frac{\langle \lambda, (m-1)\pi_{11} + (n-m)\pi_{01} -
    (n-1) \pi_{00} \rangle}{\sqrt{\langle \lambda, ((m-1) \eta_{11} +

@@

eigenvalue problem as in Theorem~\ref{thm:1}. 

@@

  \label{thm:2}

@@

\end{theorem}%
  \label{thm:3}

@@

  If $m = \Omega(\sqrt{n \log n})$, then

@@

    \frac{\Psi_{\lambda}(t^{*}) - a_{n,m}}{b_{n,m}}  \rightsquigarrow \mathcal{G}(0,1)  \end{equation}

@@

@@

\begin{IEEEproof}[Theorem \ref{thm:1}]

@@

@@

\begin{IEEEproof}[Lemma~\ref{lem:2}]

@@

\begin{IEEEproof}[Theorem~\ref{thm:2}]

@@

  \phi_{\lambda}(v;t) = \sum_{k=1}^{K} \lambda_k \sum_{\substack{uw
      \in N(v) \\ u,w \not = v}}
  \mathbb{I}\{\phi(uw,t) = k\}.
The following statements are made conditional on $|N(v)| = l$. First

@@

graph with edge probability $p_{00}$. Theorem~\ref{thm:2} then follows

@@

\begin{IEEEproof}[Theorem~\ref{thm:3}]

@@

\setminus [m]$. The following statements are made, conditional on

@@

considered in \cite{rukhin:_limit_distr_graph_scan_statis}. We have
thus provided a sketch of the arguments from
\cite{rukhin:_limit_distr_graph_scan_statis} below, along with some minor
changes to handle the case where the probability of kidney-kidney and

>>>>

\documentclass[final]{IEEEtran}
\usepackage[utopia]{mathdesign}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{mathrsfs}
\usepackage{bm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[noadjust]{cite}
%\usepackage[colon,sort&compress]{natbib}
%\numberwithin{equation}{section}
\renewcommand\arraystretch{1.2}
\let\underbrace\LaTeXunderbrace
\let\overbrace\LaTeXoverbrace
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\def\mathllap{\mathpalette\mathllapinternal}
\def\mathllapinternal#1#2{%
\llap{$\mathsurround=0pt#1{#2}$}% $
}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}
\def\mathclap{\mathpalette\mathclapinternal}
\def\mathclapinternal#1#2{%
\clap{$\mathsurround=0pt#1{#2}$}%
}
\def\mathrlap{\mathpalette\mathrlapinternal}
\def\mathrlapinternal#1#2{%
\rlap{$\mathsurround=0pt#1{#2}$}% $
}
\bibliographystyle{IEEEtran}
\begin{document}
\title{Attribute fusion in a latent process model for time series of
  graphs}
\author{Carey~E.~Priebe, Nam~H.~Lee, Youngser~Park, and Minh~Tang}%
%\thanks{Johns Hopkins University \\ Department of Applied Mathematics
%  and Statistics \\ Baltimore, Maryland 21218-2682 USA}%
\maketitle
\begin{abstract}
 We consider the problem of anomaly/change point detection given a
 time series of graphs with categorical attributes on the
 edges. Various attributed graph invariants are considered, and their
 power for detection as a function of a linear fusion parameter is
 presented.  
\end{abstract}
\begin{IEEEkeywords}
  Anomaly detection, Attributed Random Graphs, Fusion, Random Dot
  Product Graphs
\end{IEEEkeywords}
\section{Introduction}

@@

presented in

@@

      \mathbf{Q}^{(i)}(t) \equiv \mathbf{Q}_0

@@

     \bar{X}_{i}(t) \equiv \hat{\pi}_0.

@@

     \bar{X}_{i}(t) \equiv \hat{\pi}_1 \quad \text{for $i \leq m$} \\
     \bar{X}_{i}(t) \equiv \hat{\pi}_0 \quad \text{for $i > m$} 

@@

 where $\hat{\pi}_0$ and $\hat{\pi}_1$ are sub-probability vectors

@@

  \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. Define
  $\mathbf{Z}_0$ and $\mathbf{Z}_1$ by

@@

  for the continuous-time Markov chain on $\mathscr{K}_{+}$ with intensity

@@

    \Sigma_0 = \mathrm{diag}(\pi_0) \mathbf{Z}_0 +
    \mathbf{Z}_0^{T}

@@

    \Sigma_1 = \mathrm{diag}(\pi_1) \mathbf{Z}_1 +
    \mathbf{Z}_1^{T}

@@

  The second order approximation $\widehat{\mathscr{V}}$ of
  $\mathscr{V}$ is the sequence $\{\widehat{X}(t),
  \widehat{G}(t)\}_{t=1}^{\infty}$ where

@@

  \item For each $t$ and each $i \in [n]$, $\widehat{X}_i(t)$ is a

@@

    vector $Z_{i}(t)$ to $\mathscr{S}$ with mean and covariance
    matrices given below.

@@

      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_0 \\
      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_0

@@

      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_0, \quad \text{for $i \leq

@@

      \mathbb{E}[Z_i(t)] \equiv \hat{\pi}_1, \quad \text{for $i > 

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{0} \quad \text{for $i

@@

      \mathrm{Var}[Z_i(t)] \equiv \widehat{\Sigma}_{1} \quad \text{for $i

@@

    where $\widehat{\Sigma}_0$ and $\widehat{\Sigma}_1$ are the $K
    \times K$ matrices
    obtained by removing the last row and column of $\Sigma_0^{(r)}$ and
    $\Sigma_1^{(r)}$, respectively. 

@@

Suppose that we have a sequence $\{\mathscr{V}^{r} \colon r > 0 \}$ of
rdpm with vertices $[n]$ and attributes $\mathscr{K}$ where for each
$r > 0$, $\mathscr{V}^{r}$ has change parameters $(t^{*}, m, \pi_0,
\pi_1, r \mathbf{Q}_0, r \mathbf{Q}_1)$. The parameter $r$ can be
thought of as the vertex process rate, i.e., the waiting time
decreases exponentially as $r$ increases. Let us now consider the
sequence of first approximation $\bar{\mathscr{V}}^{r}$ and second
approximations $\widehat{\mathscr{V}}^{r}$ of $\mathscr{V}$. We note
that $\bar{\mathscr{V}}^{r_1} = \bar{\mathscr{V}}^{r_2}$ for any $r_1,
r_2 > 0$. Let us then denote by $\bar{\mathscr{V}}$ the (a.e.) unique
first order approximation of $\mathscr{V}^{r}$ for $r > 0$. If we
denote by $\widehat{\Sigma}_{0}^{(r)}$ and
$\widehat{\Sigma}_{1}^{(r)}$ the matrices $\widehat{\Sigma}_0$ and
$\widehat{\Sigma}_1$ for $\widehat{\mathscr{V}}^{r}$, then
$\widehat{\Sigma}_0^{(r)} = \widehat{\Sigma}_0^{(1)}/r$ and
$\widehat{\Sigma}_{1}^{(r)} = \widehat{\Sigma}_1^{(1)}/r$. Therefore,
as $r \rightarrow \infty$, $\widehat{\Sigma}_0^{(r)} \rightarrow 0$
and $\widehat{\Sigma}_1^{(r)} \rightarrow 0$ and so
$\widehat{\mathscr{V}}^{r} \rightsquigarrow
\bar{\mathscr{V}}$. This indicates that for sufficiently large $r$,
there is little statistical difference amongst the $\bar{\mathscr{V}}$,
$\mathscr{V}^{r}$ and $\widehat{\mathscr{V}}$ (\cite[Theorem
2]{lee:_laten_proces_model_time_attrib_random_graph}). $\bar{\mathscr{V}}$
will thus serve as the basis for our subsequent analysis on graph
invariants for attributed random graphs. 

@@

for $l \geq 2$. As we have mentioned earlier, we will use the first
approximation $\bar{\mathscr{V}}$ of $\mathscr{V}$ in our study of the 
asymptotic theory for these graphs invariants.   

@@

  \label{thm:9}

@@

$\mu_{\lambda}$. The following corollary extends a result in

@@

\begin{corollary}
  \label{cor:1}
  Let $\zeta$ and $\xi$ be as defined in Theorem~\ref{thm:9}. Suppose

@@

\end{corollary}

@@

$T_{\lambda}^{l}(t)$ is once again a Student t-distribution with

@@

  \label{thm:8}

@@

be $\tfrac{\rho_\lambda}{\varsigma_\lambda}$. For sufficiently large $n$
and $m$, we have for $t = t^{*}$,

@@

    \sigma_A} \\ &= (1 + o(1)) \frac{\langle \lambda, (m-1)\pi_{11} + (n-m)\pi_{01} -
    (n-1) \pi_{00} \rangle}{b_m \sqrt{\langle \lambda, ((m-1) \eta_{11} +

@@

eigenvalue problem as in Corollary~\ref{cor:1}.

@@

  \label{lem:5}

@@

\end{lemma}%
  \label{lem:6}

@@

  If $m = \Omega(\sqrt{n \log n})$ and $m = O(n^{k/(k+1)})$ for some
  $k \in \mathbb{N}$ then

@@

    \frac{\Psi_{\lambda}(t^{*}) - a_{n,m}}{b_{n,m}}  \rightsquigarrow
    \mathcal{G}(0,1)  \end{equation}
  \label{thm:6}
  Let $t \in \{l+1, \dots, t^{*}$ and $\lambda \in
  $l$, $T_{\lambda}^{l}(t)$ has approximately a
  $\mathcal{G}(\rho_{\lambda}, \varsigma_{l})$ distribution with
    \label{eq:52}
      \tfrac{\sqrt{\pi}}{6} \tfrac{a_{n,m} - a_n - b_n \gamma}{b_n} & \text{if
        $t = t^{*}$} \\
      - \tfrac{\sqrt{\pi}}{6} \gamma & \text{if $t < t^{*}$} 
    \end{cases} \\
        \tfrac{\sqrt{\pi} b_{n,m}}{6 b_n} & \text{if $t = t^{*}$} \\
        \tfrac{\sqrt{\pi}}{6} & \text{if $t < t^{*}$}

@@

The dominating factor in the power approximation for $\Psi_{\lambda}$
is $\tfrac{\rho_{\lambda}}{\varsigma_{\lambda}}$. For $t = t^{*}$ and
sufficiently large $n$ and $l$, we have
  \label{eq:53}
    \frac{\rho_{\lambda}}{\varsigma_{l}} &= \frac{a_{n,m} - a_n - b_n
      \gamma}{b_{n,m}} \\
    &= (1 + o(1)) \frac{ \langle \lambda, \xi \rangle}{\langle
      \lambda, N_{\kappa} \tfrac{\sigma_{E+F}}{\sqrt{2 \log{n}}}
      \pi_{00} \rangle}
where $\xi$ is given by
 \xi =  \Bigl(\tbinom{N_{\kappa}}{2} -
    \tbinom{N_0}{2} - \tbinom{\mu_E}{2} - \mu_E \mu_F\Bigr)\pi_{00} +
    \tbinom{\mu_{E}}{2} \pi_{11} + \mu_E \mu_F \pi_{10}  
From Eq.~\eqref{eq:53}, we see that the $\lambda$ that maximimizes
$\tfrac{\rho_\lambda}{\varsigma_{\lambda}}$ would be on the boundary,
i.e., $\lambda(k) \not = 0$ for exactly one $k$. Therefore the asymptotic theory
for scan statistics, in contrast with the other graph invariants that
were considered, indicates that there are no benefits in fusing attributes.  

@@

\begin{IEEEproof}[Corollary\ref{cor:1}]

@@

\begin{IEEEproof}[Lemma~\ref{lem:3}]

@@

\begin{IEEEproof}[Theorem~\ref{thm:8}]

@@

\begin{IEEEproof}[Lemma~\ref{lem:5}]

@@

  \phi_{\lambda}(v;t) = \sum_{\substack{uw
      \in N(v) \\ u,w \not = v}} \langle \lambda,
  \widehat{\Gamma}_{uw} \rangle
The following statements are conditional on $|N(v)| = l$. First

@@

graph with edge probability $p_{00}$. The lemma then follows

@@

\begin{IEEEproof}[Lemma~\ref{lem:6}]

@@

\setminus [m]$. The following statements are conditional on

@@

considered in \cite{rukhin:_limit_distr_graph_scan_statis}. We
provided a sketch of the arguments from
\cite{rukhin:_limit_distr_graph_scan_statis} below, along with some
minor changes to handle the case where the probability of
kidney-kidney and
